### 2024-11-29 paper/paper.qmd
initial uploads

---
title: "Predicting Goodreads Ratings Across Genres"
subtitle: "My subtitle if needed"
author: 
  - Dhruv Gupta
thanks: "Code and data are available at: https://github.com/dhruv5423/GoodReads-Cross-Genre-Predictions"
date: today
date-format: long
toc: true
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Goodreads, founded in 2006, boasts over 150 million users and is perhaps the most popular online community for4 readers. At it's heart, Goodreads is akin to a reader's social media platform, allowing them to display, discover, review, and catalog books they love (or hate) across multiple different genres.

Personal Motivation

Estimand

Results

Broader Context

# Data

## Sourcing

## Note on Measurement

## Outcome Variables

## Predictor Variables

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
## Model Set-Up

### Model Justification

# Results

# Discussion

## First Discussion Point

## Second Discussion Point

## Third Discussion Point

## Weaknesses and Next Steps

\newpage

# Appendix

# Addtional Data Details

# Model Details

## Posterior Predictive Check

## Diagnostics

\newpage

# References

### 2024-11-29 scripts/download.R

@@ -1,0 +1,95 @@
+# Load required libraries
+library(jsonlite)
+library(data.table)
+
+# Define directories
+json_dir <- "data/Raw_JSON_Files"  # Directory with the .json files
+output_dir <- "data/Sampled_JSON"  # Directory to save sampled .json files
+
+# Ensure output directory exists
+if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
+
+# Function to sample a large .json file
+sample_large_json <- function(input_file, output_file, sample_size = 1e6, chunk_size = 5000) {
+  # Open the .json file for streaming
+  con <- file(input_file, "r")
+  
+  # Initialize an empty data.table to store the sample
+  sampled_data <- NULL
+  
+  # Initialize a counter for chunks
+  chunk_count <- 0
+  
+  # Read the JSON file in chunks and randomly sample
+  while (TRUE) {
+    # Read a chunk of data
+    chunk <- tryCatch({
+      stream_in(con, pagesize = chunk_size, verbose = FALSE)
+    }, error = function(e) {
+      return(NULL)  # End of file or error
+    })
+    
+    # If no more data, break the loop
+    if (is.null(chunk) || nrow(chunk) == 0) {
+      break
+    }
+    
+    # Convert chunk to data.table
+    chunk <- as.data.table(chunk)
+    
+    # Append to sampled_data with random sampling
+    if (is.null(sampled_data)) {
+      sampled_data <- chunk[sample(.N, min(.N, sample_size))]
+    } else {
+      sampled_data <- rbind(sampled_data, chunk[sample(.N, min(.N, sample_size - nrow(sampled_data)))])
+    }
+    
+    # Increment chunk counter and print progress
+    chunk_count <- chunk_count + 1
+    cat("Processed chunk", chunk_count, "- Total sampled rows so far:", nrow(sampled_data), "\n")
+    
+    # Stop if we’ve reached the sample size
+    if (nrow(sampled_data) >= sample_size) {
+      cat("Reached sample size of", sample_size, "rows. Stopping early.\n")
+      break
+    }
+  }
+  
+  # Close the connection
+  close(con)
+  
+  # Save the sampled data as a .json file
+  write_json(as.list(sampled_data), output_file, pretty = TRUE)
+  
+  # Return the number of sampled rows
+  return(nrow(sampled_data))
+}
+
+# Define input and output files
+datasets <- list(
+  list(
[…]

### 2024-12-03 paper/paper.qmd
donwload data, raw data, and simulate data

@@
 ---
-title: "Predicting Goodreads Ratings Across Genres"
+title: "How Well Do Strike Rates Translate from T20 to ODI Formats for Top 100 vs. Non-Top 100 players?"
 subtitle: "My subtitle if needed"
@@
   - Dhruv Gupta
-thanks: "Code and data are available at: https://github.com/dhruv5423/GoodReads-Cross-Genre-Predictions"
+thanks: "Code and data are available at: https://github.com/dhruv5423/Top-100-Batters-T20-to-ODI-Strike-Rate-Mapping"
 date: today
@@
 # Introduction
-
-Goodreads, founded in 2006, boasts over 150 million users and is perhaps the most popular online community for4 readers.
-At it's heart, Goodreads is akin to a reader's social media platform, allowing them to display, discover, review, and catalog books they love (or hate) across multiple different genres.
 

### 2024-12-03 scripts/download.R

@@ -1,95 +1,42 @@
-# Load required libraries
-library(jsonlite)
-library(data.table)
+# Load required library
+if (!requireNamespace("cricketdata", quietly = TRUE)) {
+  install.packages("cricketdata")  # Install cricketdata
+}
+library(cricketdata)
 
-# Define directories
-json_dir <- "data/Raw_JSON_Files"  # Directory with the .json files
-output_dir <- "data/Sampled_JSON"  # Directory to save sampled .json files
-
-# Ensure output directory exists
-if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
-
-# Function to sample a large .json file
-sample_large_json <- function(input_file, output_file, sample_size = 1e6, chunk_size = 5000) {
-  # Open the .json file for streaming
-  con <- file(input_file, "r")
-  
-  # Initialize an empty data.table to store the sample
-  sampled_data <- NULL
-  
-  # Initialize a counter for chunks
-  chunk_count <- 0
-  
-  # Read the JSON file in chunks and randomly sample
-  while (TRUE) {
-    # Read a chunk of data
-    chunk <- tryCatch({
-      stream_in(con, pagesize = chunk_size, verbose = FALSE)
-    }, error = function(e) {
-      return(NULL)  # End of file or error
-    })
-    
-    # If no more data, break the loop
-    if (is.null(chunk) || nrow(chunk) == 0) {
-      break
-    }
-    
-    # Convert chunk to data.table
-    chunk <- as.data.table(chunk)
-    
-    # Append to sampled_data with random sampling
-    if (is.null(sampled_data)) {
-      sampled_data <- chunk[sample(.N, min(.N, sample_size))]
-    } else {
-      sampled_data <- rbind(sampled_data, chunk[sample(.N, min(.N, sample_size - nrow(sampled_data)))])
-    }
-    
-    # Increment chunk counter and print progress
-    chunk_count <- chunk_count + 1
-    cat("Processed chunk", chunk_count, "- Total sampled rows so far:", nrow(sampled_data), "\n")
-    
-    # Stop if we’ve reached the sample size
-    if (nrow(sampled_data) >= sample_size) {
-      cat("Reached sample size of", sample_size, "rows. Stopping early.\n")
-      break
-    }
-  }
-  
-  # Close the connection
-  close(con)
-  
-  # Save the sampled data as a .json file
[…]

### 2024-12-03 scripts/simulate.R

@@ -1,0 +1,167 @@
+# Load required library
+library(dplyr)
+
+# Ensure the directories exist
+simulated_data_dir <- "data/simulated_data"
+if (!dir.exists(simulated_data_dir)) {
+  dir.create(simulated_data_dir, recursive = TRUE)
+}
+
+# Function to simulate one innings of a cricket match
+simulate_innings <- function(overs, batting_team, bowling_team, players, better_players = NULL) {
+  # Create a template for ball-by-ball data
+  balls_per_over <- 6
+  total_balls <- overs * balls_per_over
+  
+  # Initialize variables
+  match_id <- 1
+  innings <- 1
+  striker <- players[1]
+  non_striker <- players[2]
+  bowler <- players[11]
+  batting_order <- players[3:length(players)]
+  current_batting_order <- 3  # Next player to come in
+  
+  # Simulate each ball
+  innings_data <- data.frame(
+    match_id = integer(total_balls),
+    innings = integer(total_balls),
+    ball = numeric(total_balls),
+    batting_team = character(total_balls),
+    bowling_team = character(total_balls),
+    striker = character(total_balls),
+    non_striker = character(total_balls),
+    bowler = character(total_balls),
+    runs_off_bat = integer(total_balls),
+    extras = integer(total_balls),
+    wides = integer(total_balls),
+    noballs = integer(total_balls),
+    byes = integer(total_balls),
+    legbyes = integer(total_balls),
+    penalty = integer(total_balls),
+    wicket_type = character(total_balls),
+    player_dismissed = character(total_balls),
+    stringsAsFactors = FALSE
+  )
+  
+  # Probabilities for "regular players"
+  regular_run_distribution <- c(0, 1, 2, 3, 4, 6)
+  regular_run_probs <- c(0.4, 0.3, 0.1, 0.05, 0.1, 0.05)
+  regular_dismissal_prob <- 0.03
+  regular_extra_prob <- 0.1
+  
+  # Probabilities for "better players"
+  better_run_distribution <- c(0, 1, 2, 3, 4, 6)
+  better_run_probs <- c(0.3, 0.3, 0.15, 0.05, 0.15, 0.05)  # Higher chance of boundaries
+  better_dismissal_prob <- 0.015  # Lower dismissal chance
+  better_extra_prob <- 0.05  # Lower chance of extras
+  
+  for (i in 1:total_balls) {
+    # Calculate ball number
+    over <- floor((i - 1) / balls_per_over) + 1
[…]

### 2024-12-03 scripts/clean.R
simulate, clean, test, model scripts; and creating model

@@ -1,0 +1,59 @@
+# Load required libraries
+library(dplyr)
+library(arrow)  # For saving and reading Parquet files
+
+# Define thresholds
+threshold_t20 <- 200   # Minimum balls faced in T20 matches
+threshold_odi <- 500   # Minimum balls faced in ODI matches
+combined_threshold <- 700  # Minimum balls faced across both formats
+
+# Step 1: Filter players meeting individual and combined thresholds
+filtered_data <- cleaned_data %>%
+  filter(
+    (match_type == "T20" & total_balls >= threshold_t20) | 
+      (match_type == "ODI" & total_balls >= threshold_odi)
+  ) %>%
+  group_by(striker) %>%
+  summarise(
+    total_balls_faced = sum(total_balls),  # Total balls faced across formats
+    .groups = "drop"
+  ) %>%
+  filter(total_balls_faced >= combined_threshold) %>%
+  inner_join(cleaned_data, by = "striker")  # Rejoin the data to include strike rates
+
+# Step 2: Calculate performance metrics for each player
+# - Include average strike rates for T20 and ODI formats
+average_strike_rate <- filtered_data %>%
+  group_by(striker) %>%
+  summarise(
+    t20_strike_rate = round(mean(strike_rate[match_type == "T20"], na.rm = TRUE), 2),  # Average T20 strike rate
+    odi_strike_rate = round(mean(strike_rate[match_type == "ODI"], na.rm = TRUE), 2),  # Average ODI strike rate
+    total_balls_faced = sum(total_balls),  # Total balls faced across formats
+    .groups = "drop"
+  )
+
+# Step 3: Rank players and assign player type
+player_rankings <- average_strike_rate %>%
+  arrange(desc((t20_strike_rate + odi_strike_rate) / 2)) %>%  # Rank by combined strike rates
+  mutate(
+    player_rank = row_number(),
+    player_type = ifelse(player_rank <= 100, "Top 100", "Non-Top 100")  # Categorize players
+  )
+
+# Step 4: Finalize the data structure
+# Merge player type information into the average_strike_rate dataset
+final_data <- average_strike_rate %>%
+  left_join(player_rankings %>% select(striker, player_type), by = "striker") %>%
+  # Remove rows with NA in either T20 or ODI strike rates
+  filter(!is.na(t20_strike_rate) & !is.na(odi_strike_rate))
+
+# Save the final dataset
+processed_data_dir <- "data/processed_data"
[…]

### 2024-12-03 scripts/download.R

@@ -1,42 +1,0 @@
-# Load required library
-if (!requireNamespace("cricketdata", quietly = TRUE)) {
-  install.packages("cricketdata")  # Install cricketdata
-}
-library(cricketdata)
-
-# Define a directory to save the raw data
-raw_data_dir <- "data/raw_data"
-if (!dir.exists(raw_data_dir)) {
-  dir.create(raw_data_dir, recursive = TRUE)
-}
-
-# Function to save raw datasets
-save_raw_data <- function(data, filename) {
-  file_path <- file.path(raw_data_dir, filename)
-  write.csv(data, file_path, row.names = FALSE)
-  cat("Saved", filename, "to", file_path, "\n")
-}
-
-# Fetch player-level data
-cat("Downloading player-level data...\n")
-player_stats <- fetch_cricsheet(type = "player", gender = "male")
-
-# Save player-level data to raw_data folder
-save_raw_data(player_stats, "player_stats.csv")
-
-# Fetch ball-by-ball data for T20 matches
-cat("Downloading T20 ball-by-ball data...\n")
-t20_bbb <- fetch_cricsheet(type = "bbb", gender = "male")
-
-# Save T20 ball-by-ball data
-save_raw_data(t20_bbb, "t20_ball_by_ball.csv")
-
-# Fetch ball-by-ball data for ODI matches
-cat("Downloading ODI ball-by-ball data...\n")
-odi_bbb <- fetch_cricsheet(type = "bbb", gender = "male")
-
-# Save ODI ball-by-ball data
-save_raw_data(odi_bbb, "odi_ball_by_ball.csv")
-
-cat("All datasets downloaded and saved to", raw_data_dir, "\n")
-

### 2024-12-03 scripts/download_combine.R

@@ -1,0 +1,55 @@
+# Load necessary library
+if (!requireNamespace("dplyr", quietly = TRUE)) {
+  install.packages("dplyr")
+}
+library(dplyr)
+
+# Define paths to the folders containing the CSV files
+odis_folder <- "data/raw_data/odis_male_csv2"
+t20s_folder <- "data/raw_data/t20s_male_csv2"
+
+# Function to read and combine all CSV files in a folder, excluding "_info" files
+combine_csv_files <- function(folder_path) {
+  # List all CSV files in the folder
+  csv_files <- list.files(folder_path, pattern = "\\.csv$", full.names = TRUE)
+  
+  # Exclude files with "_info" in their name
+  csv_files <- csv_files[!grepl("_info", csv_files)]
+  
+  # Read each CSV file and combine them into one data frame
+  combined_data <- lapply(csv_files, function(file) {
+    data <- read.csv(file, row.names = NULL, check.names = FALSE)
+    
+    # Standardize column types if necessary (e.g., season to character)
+    if ("season" %in% colnames(data)) {
+      data$season <- as.character(data$season)
+    }
+    
+    return(data)
+  }) %>%
+    bind_rows()  # Combine all data frames
+  
+  return(combined_data)
+}
+
+# Combine ODI data
+cat("Combining ODI CSV files...\n")
+odi_data <- combine_csv_files(odis_folder)
+cat("ODI data successfully combined!\n")
+
+# Combine T20 data
+cat("Combining T20 CSV files...\n")
+t20_data <- combine_csv_files(t20s_folder)
+cat("T20 data successfully combined!\n")
+
+# Ensure the processed data folder exists
+processed_data_dir <- "data/processed_data"
+if (!dir.exists(processed_data_dir)) {
+  dir.create(processed_data_dir, recursive = TRUE)
+}
+
+# Save the combined data frames as new CSV files
+write.csv(odi_data, file.path(processed_data_dir, "combined_odi_data.csv"), row.names = FALSE)
+write.csv(t20_data, file.path(processed_data_dir, "combined_t20_data.csv"), row.names = FALSE)
+
+cat("Combined datasets saved to 'data/processed_data/' folder.\n")

### 2024-12-03 scripts/model.R

@@ -1,0 +1,37 @@
+# Load necessary libraries
+library(lme4)
+library(dplyr)
+library(arrow)
+
+# Load the dataset
+data_path <- "data/processed_data/final_analysis_data.parquet"  # Adjusted to the correct path
+model_data <- read_parquet(data_path)
+
+# Ensure correct data types for categorical variables
+model_data <- model_data %>%
+  mutate(
+    player_type_binary = ifelse(player_type == "Top 100", 1, 0)  # Convert player_type to binary
+  )
+
+# Check if player has multiple observations
+player_obs_count <- model_data %>%
+  group_by(striker) %>%
+  summarise(n = n())
+
+# Simplified model without random effects if players have only one observation
+strike_rate_model <- lm(
+  formula = odi_strike_rate ~ t20_strike_rate + player_type_binary + t20_strike_rate:player_type_binary,
+  data = model_data
+)
+
+# Display the model summary
+summary(strike_rate_model)
+
+# Save the fitted model as an RDS file
+model_output_path <- "models/strike_rate_model_simple.rds"
+if (!dir.exists("models")) {
+  dir.create("models", recursive = TRUE)
+}
+saveRDS(strike_rate_model, model_output_path)
+
+cat("Simplified linear model saved to:", model_output_path, "\n")

### 2024-12-03 scripts/simulate.R

@@ -1,3 +1,6 @@
+# Set seed for reproducibility
+set.seed(123)
+
 # Load required library
 library(dplyr)
 

### 2024-12-03 scripts/test.R

@@ -1,0 +1,75 @@
+# Load required libraries
+library(testthat)
+library(dplyr)
+library(arrow)
+
+# Load the cleaned dataset (read-only)
+processed_data_dir <- "data/processed_data"
+parquet_file_path <- file.path(processed_data_dir, "final_analysis_data.parquet")
+test_data <- read_parquet(parquet_file_path)
+
+# Copy the data into a new object for testing (no changes will affect the original file)
+testing_data <- test_data
+
+# Define test thresholds
+threshold_t20 <- 200   # Minimum balls faced in T20 matches
+threshold_odi <- 500   # Minimum balls faced in ODI matches
+combined_threshold <- 700  # Minimum balls faced across both formats
+
+# Start test suite
+test_that("Final dataset structure is correct", {
+  # Check if required columns exist
+  expected_columns <- c("striker", "t20_strike_rate", "odi_strike_rate", "total_balls_faced", "player_type")
+  expect_true(all(expected_columns %in% colnames(testing_data)), "All required columns must be present")
+  
+  # Check column types using `expect_type` for numeric and character columns
+  expect_type(testing_data$t20_strike_rate, "double")
+  expect_type(testing_data$odi_strike_rate, "double")
+  expect_true(is.numeric(testing_data$total_balls_faced), "Total balls faced must be numeric (integer or double)")
+  expect_type(testing_data$player_type, "character")
+})
+
+test_that("Final dataset meets data thresholds", {
+  # Validate T20 and ODI thresholds
+  expect_true(all(testing_data$t20_strike_rate >= 0, na.rm = TRUE), "T20 strike rates must be non-negative")
+  expect_true(all(testing_data$odi_strike_rate >= 0, na.rm = TRUE), "ODI strike rates must be non-negative")
+  expect_true(all(testing_data$total_balls_faced >= combined_threshold, na.rm = TRUE), "Players must meet combined threshold for balls faced")
+})
+
+test_that("Player categorization is accurate", {
+  # Check for valid player_type values
+  expect_true(all(testing_data$player_type %in% c("Top 100", "Non-Top 100")), "Player type must be 'Top 100' or 'Non-Top 100'")
+  
+  # Validate top 100 rankings
+  top_100_players <- testing_data %>%
+    filter(player_type == "Top 100") %>%
+    arrange(desc((t20_strike_rate + odi_strike_rate) / 2)) %>%
+    mutate(player_rank = row_number())
+  
[…]

### 2024-12-04 paper/paper.qmd
simulate, clean, test, model scripts; and creating model
edits to paper, figures, references

@@
 ---
+```{r}
+#| include: false
+#| warning: false
+#| message: false
 
+#Loading Packages
+library(tidyverse)
+library(lintr)
+library(lme4)
+library(here)
+library(dplyr)
+library(arrow)
+library(knitr)
+library(tinytex)
+library(kableExtra)
+library(janitor)
+
+# Set the file path
+file_path <- here::here("data", "processed_data", "combined_odi_data.csv")
+
+# Load the data and clean column names
+combined_odi_data <- read.csv(file_path) %>%
+  clean_names()  # Standardize column names
+
+# Function to clean column names for display
+format_column_names <- function(data) {
+  colnames(data) <- gsub("_", " ", colnames(data))   # Replace underscores with spaces
+  colnames(data) <- tools::toTitleCase(colnames(data)) # Capitalize words
+  return(data)
+}
+
+# Set seed for reproducibility
+set.seed(324)
+
+# Display random rows of the data in split tables with cleaned column names
+# Match Information
+match_info <- combined_odi_data %>%
+  select(match_id, season, start_date, venue, innings, ball) %>%
+  sample_n(3) %>%
+  format_column_names()
+
+# Batting and Bowling Teams
+bat_bowl_info <- combined_odi_data %>%
+  select(batting_team, bowling_team, striker, non_striker, bowler) %>%
+  sample_n(3) %>%
+  format_column_names()
+
+# Runs and Extras
+runs_extras_info <- combined_odi_data %>%
+  select(runs_off_bat, extras, wides, noballs, byes, legbyes, penalty) %>%
+  sample_n(3) %>%
+  format_column_names()
+
+# Wicket Details
+wicket_info <- combined_odi_data %>%
+  select(wicket_type, player_dismissed, other_wicket_type, other_player_dismissed) %>%
+  sample_n(3) %>%
+  format_column_names()
+
+# Define the path to the final analysis data
+file_path <-  here::here("data", "processed_data", "final_analysis_data.parquet")
+
+# Load the final analysis data
+final_data <- read_parquet(file_path)
+
+# Load the model data
+model_data_path <- here::here("data", "processed_data", "final_analysis_data.parquet")
+model_data <- read_parquet(model_data_path)
+
+# Load the saved model
+model_path <- here::here("models", "strike_rate_model_simple.rds")
+strike_rate_model <- readRDS(model_path)
+
+```
 # Introduction
 
-Personal Motivation
+The sport of cricket has evolved greatly in recent times.
+Traditionally played over the course of 5 days, new, shorter formats are becoming increasingly popular due to their accessibility and faster pace of play.
+One Day International (ODI) and T20 formats are foremost among these formats - consisting of one innings of 50 overs and 20 overs respectively for either side.
+[@icc_three] All three formats, Test, ODI, and T20 present different challenges and require different skill-sets.
+Lasting 5 days, test matches require endurance, patience and strategy.
+In comparison, T20 cricket is much more exciting, with batsmen 'aiming for boundaries' as the teams try to score as many runs as possible in this short format.
+ODI cricket is a balance of the two - while relatively fast, strategy is key and slower play is more prevalent.
+[@behind]
+
+A batter's strike rate is a key performance indicator of their ability to score runs.
+It is defined as the 'runs per 100 balls' they face.
+[@sportmakers] For example, if a batter scores 100 runs in 50 balls, their strike rate can be calculated to be 200.
+Given the differences in strategy between formats, it is common to expect differences in strike rates.
+Longer formats generally have lower strike rates, and shorter formats higher.
+But how well do strike rates translate across formats, particularly for top-performing players compared to the rest of field?
+This is of particular interest, not only in evaluating individual player adaptability, but in understanding how skills in one format can be extrapolated to another.
+
+This paper aims to investigate whether how well strike rates translate from T20 to ODI formats, for top-performing players compared to their peers.
+Specifically, we aim to estimate the difference in ODI strike rate attributable to 1) T20 Strike Rates, 2) Player Group (whether they are a top performing batter or not), and 3) the interaction between the two factors.
+Our estimate is the conditional effect of T20 Strike Rate on ODI Strike Rate, adjusted for player type, while accounting for player-specific variability.
 
@@
 
-# Data
+The rest of paper is structured as follows..
 
-## Sourcing
+# Data {#sec-data}
 
-## Note on Measurement
+## Data Selection and Sourcing
 
-## Outcome Variables
+Data used in this paper has been sourced from the reputable cricket database, CricSheet, [@cricsheet].
+Created by Stephen Rushe, CricSheet is an open-source platform providing detail ball-by-ball data for 17,944 cricket matches.
+Contained in the database were 2,858 One Day International matches and 3,856 T20 International matches.
+This paper focuses on the male cricket matches, with 2,391 Male ODIs and 2,499 Male T20 match data downloaded as CSV files.
+This data set is particularly well-suited for this study because it ensures consistent measurement across matches, detailed coverage of player-level performance, and a large sample size, enabling robust statistical analysis of performance differences across formats.
 
-## Predictor Variables
+Additional details on the data selection process and a step-by-step guide to access and download the data are provided in [@sec-data-details].
 
-# Model
+## Raw Data
 
-The goal of our modelling strategy is twofold. Firstly,...
+Below are samples from the raw data regarding ODI matches after all raw csv files of individual match data were combined.
+The data is split into logical groups for clarity, but are all part of the same data frame.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+### Match Information
+
+@tbl-raw1 contains basic information about the match: match ID, season, start date, venue, innings, and ball.
+So, the first row would give match information of Match ID 1131279 at the ICC Academy in the 2017/2018 season, on the 1st of January 2023.
+The ball that the data captured is the 4th ball from the 44th over, in the first innings.
+
+```{r}
+#| label: tbl-raw1
+#| tbl-cap: "Raw Data - Sample Match Information"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Display match information
+knitr::kable(match_info, row.names = FALSE)
+```
+
+### Batting and Bowling Teams
+@tbl-raw2 is about the batting team, bowling team, and key players involved at the time of the delivery.
+The first row can be understood to show that the time of delivery 44.4 of Match 1131279 (correponding to row 1 in @tbl-raw1), the batting team is India, the bowling team is New Zealand, the Striker is V Kohli, the other batsman (or the non-striker) is SS Iyer, and the Bowler is TA Boult.
+```{r}
+#| label: tbl-raw2
+#| tbl-cap: "Raw Data - Sample Batting and Bowling Teams Information"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Display batting and bowling teams information
+knitr::kable(bat_bowl_info, row.names = FALSE)
+```
+
+### Runs and Extras
+@tbl-raw3 contains the runs scored, extras such as wides, no-balls, byes, leg-byes, and penalties.
+So for the same ball in @tbl-raw1 and @tbl-raw2, 1 run was scored by V Kohli.
+There were no extras, wides, no balls, byes, legbyes, or penalties.
+```{r}
+#| label: tbl-raw3
+#| tbl-cap: "Raw Data - Sample Runs and Extras Information"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Display runs and extras information
+knitr::kable(runs_extras_info, row.names = FALSE)
+```
+
+### Wicket Details
+@tbl-raw4 contains information about wickets, including the type of dismissal and the players involved. For each of the 3 balls displayed in this sample there were no wickets or players dismissed.
+```{r}
+#| label: tbl-raw4
+#| tbl-cap: "Raw Data - Sample Wicket Details"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Display wicket details information
+knitr::kable(wicket_info, row.names = FALSE)
+```
+
+## Cleaning
+
+### Desired Structure of Analysis Data
+Refer to the "Sketches" section of the repository.
+There was an understanding of how the dataset needed to be transformed in order to get it ready for analysis through our model.
+he desired structure of the data aimed to include the following key variables: player name, player type, match type, strike rate, and total balls faced.
+Below is an example of a representation of the desired structure:
+
+```{r}
+#| label: tbl-desired
+#| tbl-cap: "Desired Structure of Analysis Data"
+#| echo: false
+#| warning: false
+#| message: false
+
+library(knitr)
+library(kableExtra)
+
+# Example of the desired table structure
+desired_structure <- data.frame(
+  Player = c("Player A", "Player A", "Player B", "Player B"),
+  Player_Type = c("Top 100", "Top 100", "Non-Top 100", "Non-Top 100"),
+  Match_Type = c("T20", "ODI", "T20", "ODI"),
+  Strike_Rate = c(140.0, 90.0, 120.0, 75.0),
+  Total_Balls_Faced = c(1200, 1200, 800, 800)
+)
+
+# Clean column names to remove underscores
+colnames(desired_structure) <- gsub("_", " ", colnames(desired_structure))
+
+# Create the table
+kable(desired_structure, align = 'lccrr') %>%
+  kable_styling(full_width = FALSE, position = "center")
+```
+
+### Filtering on Balls Faced
+
+The first step after downloading raw data files and combining them into large T20 and ODI ball by ball data files was to filter players who had a large enough sample size across formats.
+That is to say, only those who meaningfully participated.
+Importantly we wanted to exclude players whose sample size of balls faced were too few, as these observations could skew the findings from the paper.
+Therefore, players were required to meet a minimum threshold of 200 balls faced in T20 matches and 500 balls faced in ODI matches.
+The combined balls faced in both formats needed to be above a threshold of 700.
+In doing so, we avoid noise from low-sample sizes.
+The library dplyr, [@R-dplyr], was used to filter and aggregate the data.
+
+### Calculating Average Strike Rates, Identifying Top 100 Players
+
+The 'mean' function was used to calculate the average T20 Strike Rate and average ODI strike rate. Missing or null values were exlcuded using 'na.rm = TRUE'.
+
+Because part of are model is concerned with differences between top performing players and their peers, we obtain the Top 100 batsmen before y ranking them based on the average of their T20 and ODI strike rates.
+Here, the 'mutate' function was used to calculate ranks and assign each player as either 'Top 100', or 'Non-Top 100'.
+
+### Structuring the Final Dataset
+
+The arrow package was used to save the cleaned data set as a Parquet file.
+[@R-arrow] The final cleaned dataset now included the columns "Player", "Player Type", "Match Type", "Strike Rate", "Total Balls Faced".
+@tbl-analysis-sample contains a sample of the first 5 rows of the final analysis data.
+
+```{r}
+#| label: tbl-analysis-sample
+#| tbl-cap: "Sample of Analysis Data"
+#| echo: false
+#| warning: false
+#| message: false
+
+library(dplyr)
+library(knitr)
+
+# Set seed for reproducibility
+set.seed(42)
+
+# Randomly sample 5 rows from the loaded data
+sample_data <- final_data %>%
+  sample_n(5)
+
+# Rename columns for clarity
+colnames(sample_data) <- c("Striker", "T20 Strike Rate", "ODI Strike Rate", "Total Balls Faced", "Player Type")
+
+# Display the table
+knitr::kable(sample_data, row.names = FALSE)
+
+```
+
+## Summary of Analaysis Data
+
+After getting the data ready for analysis, it may be useful to visualize the data before we run the model.
+@fig-bar-analysis displays the average strike rates for both match types, for both player groups.
+As expected, we find that Top 100 players have higher strike rates than Non-Top 100 players.
+
+```{r}
+#| label: fig-bar-analysis
+#| fig-cap: "T20 vs. ODI Average Strike Rates by Player Group"
+#| fig.width: 6
+#| fig.height: 4
+#| fig.align: center
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: "H"
+
+library(ggplot2)
+library(dplyr)
+
+# Calculate average strike rates for each player type and match type
+average_strike_rates <- final_data %>%
+  group_by(player_type) %>%
+  summarise(
+    T20 = mean(t20_strike_rate, na.rm = TRUE),
+    ODI = mean(odi_strike_rate, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  pivot_longer(
+    cols = c(T20, ODI),
+    names_to = "Match_Type",
+    values_to = "Average_Strike_Rate"
+  )
+
+# Create the bar chart
+strike_rate_bar_chart <- ggplot(average_strike_rates, aes(x = Match_Type, y = Average_Strike_Rate, fill = player_type)) +
+  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
+  labs(
+    x = "Match Type",
+    y = "Average Strike Rate",
+    fill = "Player Type"
+  ) +
+  theme_minimal()
+
+# Display the plot
+strike_rate_bar_chart
+
+```
+@tbl-average-strike-rate summarizes the average strike rates for each player type - numerically describing @fig-bar-analysis. 
+```{r}
+#| label: tbl-average-strike-rate
+#| tbl-cap: "Average Strike Rates by Player Type"
+#| echo: false
+#| warning: false
+#| message: false
+
+average_strike_rates <- final_data %>%
+  group_by(player_type) %>%
+  summarise(
+    `T20 Strike Rate` = mean(t20_strike_rate, na.rm = TRUE),
+    `ODI Strike Rate` = mean(odi_strike_rate, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  rename(`Player Type` = player_type)
+
+# Display the summary table
+knitr::kable(average_strike_rates, row.names = FALSE)
+
+```
+
+# Model {#sec-model}
+
 ## Model Set-Up
 
-### Model Justification
+In order to investigate how T20 strike rates translate into ODI Strike rates, we fit a linear regression model.
+Our outcome variable is $Y_i$, for which we attempt use 3 explanatory variables to fit: T20 Strike Rate - $X_{1i}$, a dummy variable for player type (whether they are Top 100 or not) - $X_{2i}$, and an interaction term between the two, $X_{1i} \cdot X_{2i}$.
 
-# Results
+The mathematical representation of the model is as follows:
 
-# Discussion
+$$
+Y_i = \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \beta_3 (X_{1i} \cdot X_{2i}) + \epsilon_i
+$$
+
+Where:
+
+- $Y_i$: The ODI strike rate of player $i$.
+
+- $X_{1i}$: The T20 strike rate of player $i$.
+
+- $X_{2i}$: A binary indicator of player type, where $X_{2i} = 1$ for "Top 100" players and $X_{2i} = 0$ for "Non-Top 100" players.
+
+- $X_{1i} \cdot X_{2i}$: The interaction term between T20 strike rate and player type.
+
+- $\beta_0$: The intercept, representing the predicted ODI strike rate for "Non-Top 100" players with a T20 strike rate of 0.
+
+- $\beta_1$: The main effect of T20 strike rate, representing the change in ODI strike rate for every 1-unit increase in T20 strike rate among "Non-Top 100"
+
+- $\beta_2$: The main effect of player type, representing the difference in ODI strike rate between "Top 100" and "Non-Top 100" players when T20 strike rate is 0.
+
+- $\beta_3$: The interaction effect, representing how the relationship between T20 and ODI strike rates differs for "Top 100" players.
+
+- $\epsilon_i$: The residual error, assumed to be normally distributed with mean 0 and constant variance $\sigma^2$.
+
+We fit the model using the `lm()` function in R, implemented in the **Base R** package.
+
+Our response variable is $Y$, ODI Strike Rate. Our predictors are the continuous variable T20 Strike Rate, $X_1$, the binary variable Player Type, $X_2$, and the interaction term $X_1 \cdot X_2$.
+
+## Model Justification
+
+We attempt to capture the relationship between T20 Strike Rates and ODI Strike Rates, while accounting for player type using a dummy variable.
+Further, the inclusion of the interaction term directly tackles the research question, allowing us to test whether the relationship between T20 and ODI Strike Rates differs between groups - demonstrating potential differences in their adaptability.
+
+In designing this model, we make the following assumptions:
+
+1.
+Linearity in Coefficients: We assume that there exists a linear relationship between T20 Strike Rates and ODI Strike Rates.
+The coefficient $\beta_1$ captures this.
+Additionally, the effect of player type, $\beta_2$, is also assumed to be linear.
+Furthermore, the coefficient of the interaction term, $\beta_3$, is also modeled linearly.
+@sec-model-details runs a diagnostic plot to check for linearity.
+
+2.
+Residuals are Normally Distributed: This is to say that the variability of ODI Strike Rate around the predicted value we fit using the model should be normally distributed, around 0.
+@sec-results contains a histogram of the residuals - the model assumes normality in the form of a bell shaped curve.
+
+3.
+Homoscedasticity: This assumption is built into the model to imply that the variance of the residuals, $\epsilon_i$, is constant across independent variables.
+That is to say that the spread of the residuals should be homogenous regardless of whether we look at the predictor $\beta1$ or $\beta2$.
+@sec-model-details runs a residuals vs.
+fitted values plot to check for homoscedasticity.
+
+4.
+No Collinearlity: This assumption states that our predictor variables are not highly correlated with each other - this can make it difficult to interpret our coefficients are we are unsure regarding the interaction between the two, and if one is influencing the other.
+
+# Results {#sec-results}
+
+## Fitted Values
+
+To understand the results of the model, we begin by examining the fitted values.
+These are the predicted ODI Strike Rates for each player based on the T20 Strike Rates, Player Type, and their interaction.
+The fitted values are computed using the following equation:
+
+$$
+\hat{y} = \beta_0 + \beta_1 \cdot \text{T20 Strike Rate} + \beta_2 \cdot \text{Player Type Binary} + \beta_3 \cdot (\text{T20 Strike Rate} \times \text{Player Type Binary})
+$$
+where:
+
+- $\hat{y}$ is the predicted ODI Strike Rate,
+
+- $\beta_0$ (Intercept) = 54.87,
+
+- $\beta_1$ (T20 Strike Rate Coefficient) = 0.19,
+
+- $\beta_2$ (Player Type Binary Coefficient) = 19.36,
+
+- $\beta_3$ (Interaction Term Coefficient) = -0.038.
+
+The fitted values provide insight into how well the model captures the observed data.
+@tbl-fitted-values shows a sample of the observed ODI Strike Rates and their corresponding fitted values, grouped by Player Type.
+
+```{r}
+#| label: tbl-fitted-values
+#| tbl-cap: "Sample of Fitted Values and Observed ODI Strike Rates"
+#| echo: false
+#| warning: false
+#| message: false
+
+library(dplyr)
+library(knitr)
+
+# Add fitted values to the data
+model_data <- model_data %>%
+  mutate(
+    Fitted_Values = predict(strike_rate_model),
+    Residuals = odi_strike_rate - Fitted_Values
+  )
+
+# Sample 10 rows for display
+fitted_values_sample <- model_data %>%
+  select(striker, player_type, t20_strike_rate, odi_strike_rate, Fitted_Values, Residuals) %>%
+  sample_n(10) %>%
+  rename(
+    "Striker" = striker,
+    "Player Type" = player_type,
+    "T20 Strike Rate" = t20_strike_rate,
+    "ODI Strike Rate (Observed)" = odi_strike_rate,
+    "ODI Strike Rate (Fitted)" = Fitted_Values,
+    "Residuals" = Residuals
+  )
+
+# Display the table
+knitr::kable(fitted_values_sample, digits = 2, row.names = FALSE)
+```
+
+## Graphical Representation
+
+### Predicted vs. Observed ODI Strike Rates
+
+```{r}
+#| label: fig-predicted-vs-observed
+#| fig-cap: "Predicted vs. Observed ODI Strike Rates"
+#| fig.width: 6
+#| fig.height: 4
+#| fig.align: center
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: "H"
+
+library(ggplot2)
+
+# Create a data frame for plotting
+model_output <- model_data %>%
+  mutate(
+    predicted_odi_strike_rate = predict(strike_rate_model),
+    residuals = residuals(strike_rate_model)
+  )
+
+# Plot Predicted vs Observed
+ggplot(model_output, aes(x = predicted_odi_strike_rate, y = odi_strike_rate)) +
+  geom_point(alpha = 0.6) +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    x = "Predicted ODI Strike Rate",
+    y = "Observed ODI Strike Rate"
+  ) +
+  theme_minimal()
+
+```
+
+### Effect of T20 Strike Rate on ODI Strike Rate by Player Type
+
+```{r}
+#| label: fig-t20-vs-odi-player-type
+#| fig-cap: "Effect of T20 Strike Rate on Predicted ODI Strike Rate by Player Type"
+#| fig.width: 6
+#| fig.height: 4
+#| fig.align: center
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: "H"
+
+# Ensure correct data types
+model_output <- model_data %>%
+  mutate(
+    predicted_odi_strike_rate = predict(strike_rate_model),
+    player_type_binary = factor(player_type_binary, levels = c(0, 1), labels = c("Non-Top 100", "Top 100")),
+    t20_strike_rate = as.numeric(t20_strike_rate),
+    predicted_odi_strike_rate = as.numeric(predicted_odi_strike_rate)
+  )
+
+# Debugging: Check the structure of the model_output
+str(model_output)
+
+# Plot the effect of T20 Strike Rates by Player Type
+ggplot(model_output, aes(x = t20_strike_rate, y = predicted_odi_strike_rate, color = factor(player_type_binary))) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm", se = FALSE) +
+  scale_color_manual(values = c("blue", "orange"), labels = c("Non-Top 100", "Top 100")) +
+  labs(
+    x = "T20 Strike Rate",
+    y = "Predicted ODI Strike Rate",
+    color = "Player Type"
+  ) +
+  theme_minimal()
+
+```
+
+### Residuals Distribution
+```{r}
+#| label: fig-residuals-distribution
+#| fig-cap: "Distribution of Residuals"
+#| fig.width: 6
+#| fig.height: 4
+#| fig.align: center
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: "H"
+
+# Plot histogram of residuals
+ggplot(model_output, aes(x = residuals)) +
+  geom_histogram(binwidth = 2, fill = "skyblue", color = "black", alpha = 0.7) +
+  labs(
+    x = "Residuals",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+```
+# Discussion {#sec-discussion}
 
@@
 
-# Appendix
+# Appendix {#sec-appendix}
 
-# Addtional Data Details
+# Addtional Data Details {#sec-data-details}
 
-# Model Details
+# Model Details {#sec-model-details}
 
@@
 
+Residuals vs Fitted Values
 \newpage

### 2024-12-04 other/datasheet/datasheet_template.qmd
final changes to paper, sketches, datasheet,

---
title: "Datasheet for CricSheet's T20 and ODI Datasets"
author: 
  - Dhruv Gupta
date: today
date-format: long
abstract: "This datasheet contains information about Stephen Rushe's datasets and his website, CricSheet, which he has created"
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets. The data used is from CricSheet, an open-source platform created by Stephen Rushe [@cricsheet]

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to support cricket analytics and research, specifically to enable detailed analysis of cricket matches. It provides granular ball-by-ball data for T20 and ODI matches, allowing for performance modeling, strategy analysis, and player comparison.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by Stephen Rushe and is made available through the CricSheet platform.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The dataset is publicly available and maintained by Stephen Rushe on a volunteer basis, with some support from donations.
4. *Any other comments?*
    - The dataset has become a critical resource for researchers and cricket enthusiasts, given its high level of detail and accessibility.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance represents a ball-by-ball event in a cricket match, capturing details such as the batsman, bowler, runs scored, extras, and type of dismissal.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset contains 17,944 cricket matches, including 2,499 T20 Internationals and 2,391 ODIs, with millions of individual ball events.

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a comprehensive record of all publicly available international cricket matches but does not include all domestic or non-international games.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance includes "raw" ball-by-ball data, capturing attributes such as match ID, over and ball number, batting team, bowler, batsman, runs, extras, and dismissal type.

5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- No explicit labels are included, but data attributes such as runs and dismissals provide context for outcome modeling.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Data for some older matches might be incomplete, particularly regarding player details or venue information.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, instances are linked within a match via match ID and innings number.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No explicit splits are recommended, but users often create splits by match type (ODI vs. T20) or time period.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Minor inconsistencies or missing data may exist for older matches due to incomplete record-keeping.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Yes, the dataset is self-contained and does not rely on external resources.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the dataset consists of publicly available information.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No, the data is focused on professional cricket and does not include sensitive personal information.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, it identifies players, teams, and match types.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- Yes, players are identifiable, but the data is already public.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The dataset is designed to be open and accessible for a wide range of analytical uses.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was sourced from official match scorecards and ball-by-ball commentary.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Data was collected manually and through automated scraping tools, validated against official records.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset includes all publicly available international matches.

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The dataset is curated and maintained by Stephen Rushe.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data spans decades of cricket history, with ongoing updates.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not applicable, as the data is publicly available.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- No, the data was compiled from official scorecards.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Not applicable, as the data is public.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Not applicable, as the data is public.

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Not applicable, as the data is public.

11. *Any other comments?*
	- The dataset's reliability depends on the accuracy of the source records.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes, the dataset has been used for various cricket-related analyses, including player performance evaluation, predictive modeling of match outcomes, comparative analysis between match formats (e.g., ODI vs. T20), and strategy planning for teams.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- While there is no centralized repository for all works using the dataset, individual academic papers, blogs, and analyses frequently reference CricSheet. Examples include cricket analytics blogs and statistical analyses published by enthusiasts and researchers.

3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used for a wide variety of cricket-related tasks, including:
    - Forecasting player and team performance.
    - Identifying trends in gameplay over time.
    - Strategy optimization for teams and players.
    - Fan engagement through visualizations and interactive tools.
    - Developing machine learning models for predictive analytics in cricket.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The dataset does not contain sensitive or personal information but relies on the accuracy of publicly available data. Inaccuracies in source data could propagate into analyses. Additionally, care must be taken when applying findings to avoid perpetuating biases in player evaluation or stereotyping based on performance in specific contexts.

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for unethical purposes such as match-fixing or gambling-related activities. Additionally, it should not be used to make decisions that could negatively impact individuals without proper validation of findings.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, it is openly available on the CricSheet website.

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Via CSV files on the CricSheet website.

3. *When will the dataset be distributed?*
	- The dataset is continually updated and is currently available.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Open license, as specified by CricSheet.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Stephen Rushe.

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Contact information is available on the CricSheet website. His Mastodon server and social page can be found at https://social.deeden.co.uk/@srushe

3. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Yes, it is updated regularly with new match data.

4 *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older data remains available as part of the dataset.

\newpage

# References

### 2024-12-04 paper/paper.qmd
final changes to paper, sketches, datasheet,

@@
 ---
-title: "How Well Do Strike Rates Translate from T20 to ODI Formats for Top 100 vs. Non-Top 100 players?"
-subtitle: "My subtitle if needed"
+title: "Batting Performance Across Formats: Predicting ODI Strike Rates Using T20 Metrics"
+subtitle: "Top 100 Players Excel in Both Formats, but Adaptability Remains Uniform Across Groups"
 author: 
@@
 toc: true
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates the extent to which batting strike rates in T20 cricket predict performance in the ODI format, focusing on differences between top 100 players and their peers.
+Using ball-by-ball data from 17,944 cricket matches, we construct a linear model to estimate ODI strike rates based on T20 performance, player type, and their interaction.
+Results reveal a positive relationship between T20 and ODI strike rates for all players, with top 100 players exhibiting consistently higher ODI rates.
+Surprisingly, the interaction term is not significant, suggesting that adaptability between formats may not differ substantially between player groups.
+These findings provide a clearer understanding of performance transferability across cricket formats and highlight the nuanced factors influencing player success."
 format: pdf
 number-sections: true
-bibliography: references.bib
+bibliography: paper_references.bib
 ---
@@
 
-This paper aims to investigate whether how well strike rates translate from T20 to ODI formats, for top-performing players compared to their peers.
-Specifically, we aim to estimate the difference in ODI strike rate attributable to 1) T20 Strike Rates, 2) Player Group (whether they are a top performing batter or not), and 3) the interaction between the two factors.
-Our estimate is the conditional effect of T20 Strike Rate on ODI Strike Rate, adjusted for player type, while accounting for player-specific variability.
+In this paper, we examine how T20 batting strike rates predict ODI performance, focusing on differences between top-performing players and their peers.
+We model ODI strike rates using T20 strike rates, player type (top 100 or non-top 100), and their interaction, analyzing ball-by-ball data from nearly 17,944 matches.
+Our findings reveal that while T20 strike rates positively correlate with ODI performance, the interaction term suggests that adaptability does not differ significantly between top 100 players and others.
+This challenges the assumption that top players are uniquely skilled at transferring their performance between formats.
 
-Estimand
-
-Results
-
-Broader Context
-
-The rest of paper is structured as follows..
+These findings are important for talent evaluation and team selection strategies in cricket.
+By quantifying the relationship between T20 and ODI performance, this research provides clarity on the dynamics of skill transferability.
+The rest of paper is structured as follows: @sec-data contains information regarding data selection and sourcing, samples of the raw data, a description of the cleaning process, and an example of the cleaned analysis data.
+@sec-model contains the model used in this paper - the set up, the estimator, and the justification and assumptions used to design it.
+@sec-results provides a summary of the regression output, fitted values for the model, and graphical representation of the outcome variable.
+@sec-discussion is a discussion, analyzing the coefficients the model predicts, intepreting the graphs in @sec-results, and using literature to draw broader conclusions regarding our findings.
+It also discusses key limitations in the model.
+@sec-appendix includes details on how to download the data and challenges with regards to sampling and observing cricket data.
 
 # Data {#sec-data}
+
+The statistical software R, [@citeR], is used throughout the paper to download, clean, and analyse the data.
+Additionally, many packages within the Tidyverse, [@tidyverse], Arrow, [@R-arrow], and KableExtra, [@R-kableExtra] are used.
 
@@
 
-Additional details on the data selection process and a step-by-step guide to access and download the data are provided in [@sec-data-details].
+Additional details on the data selection process and a step-by-step guide to access and download the data are provided in the appendix [@sec-appendix].
+
+## Measurement
+
+The dataset used in this research is an organised and quantitative attempt to capture real-world cricketing phenomena, specifically player performance in the T20 and ODI formats.
+The data on CricSheet is derived from genuine ball-by-ball records of cricket matches, which are meticulously documented by official scorers throughout games.
+Each element in the dataset represents an observable event from a match, such as a ball delivered, runs scored, or wicket taken.
+This is a detailed, play-by-play breakdown of the game.
+
+The process of converting these real-world cricketing occurrences into dataset entries begins with match officials and scorers recording the results of each ball.
+CricSheet standardised and digitised these logs into CSV files.
+Key variables including strike rate, player type, and total balls faced, were derived directly from these raw observations.
+For example, strike rate is calculated as the ratio of runs scored to balls faced, multiplied by 100—a straightforward transformation of raw data into a meaningful metric.
 
@@
 
-## Model Set-Up
+## Model Set-Up {#sec-model-setup}
 
@@
 Furthermore, the coefficient of the interaction term, $\beta_3$, is also modeled linearly.
-@sec-model-details runs a diagnostic plot to check for linearity.
 
@@
 That is to say that the spread of the residuals should be homogenous regardless of whether we look at the predictor $\beta1$ or $\beta2$.
-@sec-model-details runs a residuals vs.
-fitted values plot to check for homoscedasticity.
 
@@
 # Results {#sec-results}
+
+## Regression Summary
+
+@tbl-regression-summary presents the coefficients from the linear regression model.  
+```{r}
+#| label: tbl-regression-summary
+#| tbl-cap: "Regression Coefficients for the Model"
+#| echo: false
+#| warning: false
+#| message: false
+
+library(broom)  # For cleaning regression results
+
+# Extract regression summary
+regression_summary <- broom::tidy(strike_rate_model)
+
+# Rename columns for clarity
+colnames(regression_summary) <- c("Term", "Estimate", "Standard Error", "t-value", "p-value")
+
+# Display the regression summary as a table
+knitr::kable(regression_summary)
+```
 
@@
 
-# Sample 10 rows for display
+# Sample 5 rows for display
 fitted_values_sample <- model_data %>%
   select(striker, player_type, t20_strike_rate, odi_strike_rate, Fitted_Values, Residuals) %>%
-  sample_n(10) %>%
+  sample_n(5) %>%
   rename(
@@
 
+@fig-predicted-vs-observed evaluates the accuracy of the model by comparing the predicted ODI Strike Rates to the actual observed values.  
 ```{r}
@@
 
-### Effect of T20 Strike Rate on ODI Strike Rate by Player Type
-
-```{r}
-#| label: fig-t20-vs-odi-player-type
-#| fig-cap: "Effect of T20 Strike Rate on Predicted ODI Strike Rate by Player Type"
-#| fig.width: 6
-#| fig.height: 4
-#| fig.align: center
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-pos: "H"
-
-# Ensure correct data types
-model_output <- model_data %>%
-  mutate(
-    predicted_odi_strike_rate = predict(strike_rate_model),
-    player_type_binary = factor(player_type_binary, levels = c(0, 1), labels = c("Non-Top 100", "Top 100")),
-    t20_strike_rate = as.numeric(t20_strike_rate),
-    predicted_odi_strike_rate = as.numeric(predicted_odi_strike_rate)
-  )
-
-# Debugging: Check the structure of the model_output
-str(model_output)
-
-# Plot the effect of T20 Strike Rates by Player Type
-ggplot(model_output, aes(x = t20_strike_rate, y = predicted_odi_strike_rate, color = factor(player_type_binary))) +
-  geom_point(alpha = 0.6) +
-  geom_smooth(method = "lm", se = FALSE) +
-  scale_color_manual(values = c("blue", "orange"), labels = c("Non-Top 100", "Top 100")) +
-  labs(
-    x = "T20 Strike Rate",
-    y = "Predicted ODI Strike Rate",
-    color = "Player Type"
-  ) +
-  theme_minimal()
-
-```
+\newpage
 
 ### Residuals Distribution
+
+@fig-residuals-distribution assesses the assumption of normally distributed residuals by visualizing the spread of residuals around zero, in the form of a histogram.
 ```{r}
@@
 
-## First Discussion Point
+## Intepretation of Model Coefficients
 
-## Second Discussion Point
+The linear regression model estimates the relationship between T20 and ODI strike rates while accounting for player type.
+From @tbl-regression-summary, we can use the coefficients from our model to generate the following:
 
-## Third Discussion Point
+$$
+\hat{y} = 54.87+ 0.19\cdot \text{T20 Strike Rate} + 19.36\cdot \text{Player Type Binary} - 0.037 \cdot (\text{T20 Strike Rate} \times \text{Player Type Binary})
+$$
 
-## Weaknesses and Next Steps
+The intercept, $\beta_0$, indicates that the expected strike rate for a non-top 100 player in an ODI match is 54.87.
+That is to say, if faced with 100 balls, the average non-top 100 player scores almost 55 runs.
+The coefficient $\beta_1$ = 0.19 suggests that, on average, a one-unit increase in T20 strike rate is associated with a 0.19-unit increase in ODI strike rate for non-top players, statistically significant at p = 0.002.
+This positive relationship highlights that higher T20 strike rates generally correspond to better performance in ODIs for these players.
 
+Interestingly, we observe that our dummy variable, 'Player Type Binary', has a coefficient $\beta_2$ = 19.36.
+This reflects the average difference in ODI strike rates between top and non-top players, controlling for T20 strike rate.
+Although this difference is not statistically significant (p = 0.185), it provides an approximate measure of the expected performance gap.
+As expected, higher skilled players tend to score runs at a higher rate.
+
+The interaction term has a coefficient $\beta_3$ = -0.037, which is not statistically significant (p = 0.737).
+From the predictions of the models we can reasonably say that there is no evidence that the relationship between T20 strike rates and ODI strike rates differs between top performing players and others.
+According to our model, what is a better predictor of success is the skill level that top performing batsmen exhibit.
+
+## Intepretation of Graphs
+
+@fig-predicted-vs-observed shows that most data points are clustered around the 45 degree line.
+This demonstrates that the model captures much of the variability in ODI strike rates.
+There are however, some outliers that might indicate endogeneity or other confounding variables.
+@fig-residuals-distribution show that the value of our residuals are roughly normally distributed, and centered around zero.
+This is in line with Assumption 2 in @sec-model-setup, and we can be reasonably confident that our model predictions are unbiased on average.
+However, there are some outliers towards the tails of the curve, potentially indicating the model's limitation in capturing these extreme values.
+
+## Broader Context
+
+The analysis done in this paper highlights the transferability between T20 performance and ODI performance.
+With a high degree of confidence we can say that higher T20 strike rates will generally result in high ODI strike rates, for all players - regardless of whether they are a top performing player or not.
+As we expect, top players exhibit a much higher strike rate than non top 100 players in ODI matches - but the lack of a significant interaction term tells us that our hypothesis was misguided.
+The model tells us that unintuitively, top players may not show any additional 'adaptability' separating them from their peers.
+Additionally, the low value of the coefficient $\beta_1$ indicates that it is quite difficult to increase ODI strike rate.
+For a given batsman's strike rate to increase by 1 run per 100 balls, the model predicts that they would need more than 5 runs per 100 balls in the T20 format.
+
+The findings of this study align with existing literature that emphasizes the interconnectedness of cricket formats and the transferability of batting skills.
+By highlighting the positive correlation between T20 and ODI strike rates, this research contributes to a deeper understanding of how modern batting techniques and strategies are evolving in response to the demands of different formats.
+
+## Limitations
+
+There are a fair amount limitations that are present in this model.
+First and foremost, the model uses strike rate as the sole predictor of batting performance - a top 100 player was defined purely off of strike rate.
+This approach ignores other critical aspects of batting performance, such as average runs scored, consistency, and the ability to perform under pressure.
+Players who may not have exceptionally high strike rates but consistently score runs over long periods could also be considered top performers in broader contexts.
+Thus, relying solely on strike rates could lead to an incomplete characterization of player quality.
+
+Furthermore, the model assumes linearity in the relationship between T20 and ODI strike rates, which might oversimplify the actual dynamics.
+Batting performance could exhibit non-linear trends, especially as players adapt their approach to match contexts or as they age and gain experience.
+These dynamics are not captured in the current framework
+
+There are more factors to consider.
+Environmntal and equipment constraints, such as pitch conditions, ball type, and equipment modifications, also play crucial roles in batting performance.
+These factors can affect a player's ability to adapt their skills across different formats and conditions.
+[@sage] A better model would likely incorporate these influences.
 \newpage
@@
 
-# Addtional Data Details {#sec-data-details}
+This section contains additional data details.
 
-# Model Details {#sec-model-details}
+## How to Download the Data
+To replicate the analysis, follow these steps to download the datasets:
 
-## Posterior Predictive Check
+1. Navigate to CricSheet: Visit the CricSheet website.
+2. Locate Match Type Data:
+3. Click on the relevant format (e.g., ODI, T20).
+4. Download the ball-by-ball CSV files for each format under "Men’s Cricket" section.
+5. Extract the Files:
+6. The downloaded files will be in ZIP format.
+7. Extract all files into a folder labeled according to the match format (e.g., ODI_matches/, T20_matches/).
 
-## Diagnostics
+## Sampling and Observational Data:
 
-Residuals vs Fitted Values
-\newpage
+The analysis is based on data that is observational in nature - we did not conduct experiments or control any variables.
+Observational data in sports analytics often arises from recorded performance metrics rather than randomized experiments.
+While this approach provides real-world information, it comes with limitations such as unmeasured confounding variables and selection bias.
 
-# References
+Specific to cricket there may be challenges in samping:
+
+- Inconsistent Match Representation: Players may participate in varying numbers of matches across formats, leading to unbalanced observations.
+- Player Inclusion Criteria: A threshold for balls faced was used to filter players with sufficient data (e.g., 200 balls for T20, 500 for ODI).
+This threshold minimizes noise but introduces potential bias by excluding players with fewer matches.
+- Team and Venue Effects: Factors like team strength or venue conditions could influence performance, but these are not explicitly modeled.
+
+## Idealized Methodology
+
+To address these challenges in future work, an idealized methodology could involve:
+
+- Stratified Sampling: Stratify players by their participation levels (e.g., high, medium, low match participation). This also ensures balanced representation across player groups for analysis.
+
+- Better Simulated Data: Simulate player performances under controlled conditions to test hypotheses.
+For example, simulate strike rates by varying contextual factors like opposition quality or pitch conditions.
+
+- Incorporating Hierarchical Models:Use hierarchical (mixed-effects) models to account for team and venue effects.
+This approach would allow for varying intercepts/slopes across teams or venues, reducing bias from unmeasured variables.
+
+# Citations
